---
title: "Bayesian GLM"
author: "Linda Tang"
date: "9/6/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
library(tidyverse)
library(patchwork)
library(knitr)
library(bayesplot)
library(GGally)
require(rstanarm)
require(loo)
require(bayesplot)
require(caret)
library(rstan)
library(bridgesampling)

bike <- read_csv("210830_bikecrash.csv")
```

### Question Formulation: 

**Bayesian Logistic regression**: The response is binary (high crash vs. not high crash). The response will be denoted as 1 if it has > 60 crashes per 100,000 residents per year. 

**Sampling Model**: $Y \mid \theta \sim Bern(\theta)$. We chose the logit link function $g(\theta) = log(\frac{\theta}{1-\theta}) = \eta$ and the systematic component $\eta = x^T\beta$. Overall the likelihood of a single observation is :

$$ p (Y\mid\theta) = {n \choose y}\theta^y(1-\theta)^{n-y} = {n \choose y} (\frac{e^\eta}{1+e^\eta})^y(\frac{1}{1+e^\eta})^{n-y}$$
**Prior**: Since we have little prior belief about which predictor will be significant and the relative magnitude/direction of their influence, we want to chose a weakly informative prior. We will use a t-prior for all coefficients $\beta_k$ and the intercept $\beta_o$. 

$$
\beta_1...\beta_N, \beta_o \sim \text{t} (7,0)
$$

### Data Manipulation

I created a binary response variable. I noticed that the population and pct_rural are on quite different scales, so I scaled them and saved it into the original dataframe. 

```{r}
bike <- bike %>%
  mutate(high_crashes = ifelse(crashes*100000/pop > 60, 1, 0)) %>%
  mutate(pop = scale(pop)) %>%
  mutate(pct_rural = scale(pct_rural))
```

### Fitting the model

```{r}
model <- stan_glm(high_crashes ~ pop + pct_rural, # only used these 2 predictors
                 data = bike,
                 family = binomial(link = "logit"),
                 prior = student_t(df = 7, 0),
                 prior_intercept = student_t(df = 7, 0), 
                 cores = 4,
                 chains = 4, # run 4 MCMC chains
                 iter = 1000, 
                 seed = 12345)
```

### Model Diagnoistics

Some model diagnostic plots are shown below:

```{r fig.height=5, fig.width=10}
# all chains converged & mixed well
color_scheme_set("mix-blue-red")
plot(model, "trace")
```

```{r fig.height=5, fig.width=10}
# all chains converged & mixed well
plot(model, "trace", pars = "pct_rural", ylab = "Asdf") +
  labs(title = "Trace plot: percent rural", 
       y = "Estimate", x = "Draw")
```

```{r fig.height=5, fig.width=10}
# density from the 4 chain seems to overlap well
# suggest convergence and good mixing 
plot(model, "dens_overlay")
```

```{r fig.height=5, fig.width=10}
# posterior predictive checks
# draws from the posterior predictive dist. seems to be similar to actual data
pp_check(model, plotfun = "hist", nreps = 5) + 
  xlab("Crashes") 
```


```{r fig.height=5, fig.width=10}
# posterior predictive checks
# draws from the posterior predictive dist. seems to be similar to actual data
# although there's one outlier?
pp_check(model) + 
  xlab("Crashes")
```


```{r fig.height=5, fig.width=10}
# relatively low autocorrelation 
plot(model, "acf_bar") +  
  labs(title = "ACF plots")
```

```{r fig.height=5, fig.width=10}
# Compare prior and posterior dist.
# Seems like the posterior distribution is much narrower
prior_summary(model)
posterior_vs_prior(model) 
```

### Inference

Output of model coefficients: 

```{r}
summary(model)
```

A more precise output of model coefficients: 

```{r}
round(posterior_interval(model, prob = 0.95), 3)
```

Based on the model output above, the 95% confidence interval of the coefficient of pct_rural is from -1.387 to - 0.093, which doesn't cross 0. So after accounting for population, pct_rural is still important in predicting whether a county is high crash.

Interpretation: for one standard deviation increase in pct_rural (which is around 28%), the odds of a county being high crash by multiply be a factor of 0.4965853, holding population constant. 

(*Note: Initially I build the modeling without scaling the variables and I found pct_rural to be not significant after accounting for population. However, after I scaled the data, pct_rural is significant). 

