---
title: "Survival Analysis"
author: "Linda Tang"
date: "9/27/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(survival)
library(survminer)
library(knitr)
got <- read_csv("210927_got.csv") %>%
  select(character_name, duration_in_episodes, is_dead, royal, house, gender)
```

## KM Curve Analysis

*duration_in_episodes*: Survival for characters

*is_dead*: 1, 0 indicator of death or not

*royal, house, gender*: Stratas to explore

```{r}
ggsurvplot(survfit(Surv(duration_in_episodes, is_dead) ~ 1, data = got), 
     xlab = "Episodes", ylab = "Est. Survival Probability",
     conf.int = F)
```

```{r}
ggsurvplot(survfit(Surv(duration_in_episodes, is_dead) ~ royal, 
                   data = got), 
     xlab = "Episodes", ylab = "Est. Survival Probability",
     ylim = c(0, 1),
     conf.int = T, censor = F, 
     legend.labs = c("Not Royal", "Royal"))
```

```{r}
survdiff(Surv(duration_in_episodes, is_dead) ~ royal, data = got)
```
p_value = 0.7, there's no statistically significant difference. 

```{r}
ggsurvplot(survfit(Surv(duration_in_episodes, is_dead) ~ gender, 
                   data = got), 
     xlab = "Episodes", ylab = "Est. Survival Probability",
     ylim = c(0, 1),
     conf.int = T, censor = F, 
     legend.labs = c("Female", "Male"))
```
```{r}
survdiff(Surv(duration_in_episodes, is_dead) ~ gender, data = got)
```
p_value = 0.02, there is a statistically significant difference. 

```{r}
survdiff(Surv(duration_in_episodes, is_dead) ~ house, data = got)
```

p_value = 0.009 , there's statistical evidence that the house has an impact on survival of characters (at least one house is different). 

## Modeling

*royal, house, gender*: confounding to control 

#### Exponential Model 

AFT model:

```{r}
aft_exp <- survreg(Surv(duration_in_episodes, is_dead) ~ gender + royal + house, 
                 data = got, dist = "exponential")
summary(aft_exp)
```

The p_value for I(gender = male) is 0.02366, which means it's a statistically significant predictor. Holding all else equal, a male is expected to survive approximately 0.41 times longer than a female. The probability that males survive to time 0.41t is the same females surviving to 1t. 

```{r}
resids <- (log(got$duration_in_episodes) - aft_exp$linear.predictors) / 
  (aft_exp$scale)

m1 <- survfit(Surv(resids, is_dead) ~ 1, data = got)
plot(m1, xlab = "AFT Residuals (Exponential Model)", 
     ylab = "Survival Probability")

exp.x <- seq(min(resids), max(resids), length = 100)
exp.y <- pexp(exp.x, lower.tail = F) # F(t)
lines(exp.x, exp.y, col = "red", lwd = 2)
```

The residuals doesn't overlap well with the survival function of the assumed distribution, so the Exponential model doesn't fit very well. 

#### Weibull Model

AFT model:

```{r}
aft_w <- survreg(Surv(duration_in_episodes, is_dead) ~ gender + royal + house, 
                 data = got, dist = "weibull")
summary(aft_w)
```

```{r}
resids <- (log(got$duration_in_episodes) - aft_w$linear.predictors) / 
  (aft_w$scale)

m1 <- survfit(Surv(resids, is_dead) ~ 1, data = got)
plot(m1, xlab = "AFT Residuals (Weibull)", 
     ylab = "Survival Probability")

exp.x <- seq(min(resids), max(resids), length = 100)
exp.y <- pweibull(exp.x, shape = 1/aft_w$scale, lower.tail = F) # F(t)
lines(exp.x, exp.y, col = "red", lwd = 2)
```

The residuals doesn't overlap well with the survival function of the assumed distribution, so the Weibull model doesn't fit very well. 

#### Bayesian Survival Analysis 


```{r}
install.packages("rstanarm", 
                 repos = c("https://mc-stan.org/r-packages/", 
                           getOption("repos")))
```


```{r}
library(rstanarm)
m1 <- stan_surv(Surv(duration_in_episodes, is_dead) ~ gender + 
                  royal + house, 
               data = got, basehaz = "weibull",
               prior = normal(autoscale = T),
               prior_intercept = normal(autoscale = T),
               prior_aux = cauchy(0, 5, autoscale = T),
               chains = 2, iter = 2000, seed = 123,
               prior_PD = F)
```

```{r}
summary(m1)
```
```{r}
plot(m1, plotfun = "basehaz")
```
```{r}
ps_check(m1)
```

```{r}
plot(m1, "dens_overlay")
```

```{r}
plot(m1, "trace")
```

```{r}
# m2 <- stan_surv(Surv(DWHFDAYS, DWHF) ~ TRTMT + EJF_PER + PREVMI, 
#                data = dig, basehaz = "weibull-aft",
#                prior = normal(autoscale = T),
#                prior_intercept = normal(autoscale = T),
#                prior_aux = cauchy(0, 5, autoscale = T),
#                chains = 2, iter = 2000, seed = 123,
#                prior_PD = F)
```

## Results

```{r}
round(as.data.frame(summary(m1))[,c(1, 3, 7)], 3)
```

```{r}
# round(as.data.frame(summary(m2))[,c(1, 3, 7)], 3)
```





